rgbif: a package for working with species occurrence data in R

### 2015-10-03 manuscript.Rmd
remove some things, remove not needed refs, updated circleci badge

```{r compile-settings, include=FALSE}
library("methods")
library("knitr")
opts_chunk$set(tidy = FALSE, warning = FALSE, message = FALSE,
               cache = 1, comment = NA, verbose = TRUE)
basename <- gsub(".Rmd", "", knitr:::knit_concord$get('infile'))
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))

```

@@

Users of the popular statistical and mathematical computing platform
R [@R] enjoy a wealth of readily installable comparative phylogenetic
methods and tools [@taskview]. Exploiting the opportunities arising from
this wealth for complex and integrative comparative research questions
relies on the ability to reuse and integrate previously generated or
published data and metadata. The expanding data exchange needs of the
evolutionary research community are rapidly outpacing the capabilities of
most current and widely used data exchange standards [@Vos_2012], which
were all developed a decade or more ago. This has resulted in a radiation
of different data representations and exchange standard "flavors" that are
no longer interoperable at the very time when the growth of available data
and methods has made that interoperability most valuable. In response to
the unmet needs for standardized data exchange in phylogenetics, a modern
XML-based exchange standard, called NeXML, has recently been developed
[@Vos_2012]. NeXML comprehensively supports current data exchange needs,
is predictably machine-readable, and is forward compatible.

The exchange problem for phylogenetic data is particularly acute in light
of the challenges in finding and sharing phylogenetic data without the
otherwise common loss of most data and metadata semantics [@Drew_2013;
@Stoltzfus_2012; @Cranston_2014]. For example, the still popular NEXUS
file format [@Maddison_1997] cannot consistently represent horizontal
gene transfer or ambiguity in reading a character (such as a DNA sequence
base pair).  This and other limitations have led to modifications of NEXUS
in different ways for different needs, with the unfortunate result that
NEXUS files generated by one program can be incompatible with another
[@Vos_2012]. Without a formal grammar, software based on NEXUS files
may also make inconsistent assumptions about tokens, quoting, or element
lengths.  @Vos_2012 estimates that as many as 15% of the NEXUS files in
the CIPRES portal contain unrecoverable but hard to diagnose errors.

A detailed account of how the NeXML standard addresses these and
other relevant challenges can be found in @Vos_2012. In brief,
NeXML was designed with the following important properties. First,
NeXML is defined by a precise grammar that can be programmatically
**validated**; i.e., it can be verified whether a file precisely
follows this grammar, and therefore whether it can be read (parsed)
without errors by software that uses the NeXML grammar (e.g. RNeXML)
is predictable.
Second, NeXML is **extensible**: a user
can define representations of new, previously unanticipated information
(as we will illustrate) without violating its defining grammar. Third
and most importantly, NeXML is rich in **computable semantics**: it is
designed for expressing metadata such that machines can understand
their meaning and make inferences from it. For example, OTUs in a tree
or character matrix for frog species can be linked to concepts in a
formally defined hierarchy of taxonomic concepts such as the
Vertebrate Taxonomy Ontology [@Midford2013], which enables a machine
to infer that a query for amphibia is to include the frog data in what
is returned. (For a more broader discussion of the value of such
capabilities for evolutionary and biodiversity science we refer the
reader to @Parr2011.)

@@

All software, scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI corresponds to a snapshot of the GitHub repository at [github.com/ropensci/rgbif](https://github.com/ropensci/rgbif).

>>>>

```{r compile-settings, include=FALSE}
library("methods")
library("knitr")
opts_chunk$set(tidy = FALSE, warning = FALSE, message = FALSE,
               cache = 1, comment = NA, verbose = TRUE)
basename <- gsub(".Rmd", "", knitr:::knit_concord$get('infile'))
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
```

@@

Users of the popular statistical and mathematical computing platform
R [@R] enjoy a wealth of readily installable comparative phylogenetic
methods and tools.

@@

All software, scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI corresponds to a snapshot of the GitHub repository at [github.com/sckott/msrgbif](https://github.com/sckott/msrgbif).
